---
title: "Customer Acquisition vs. Customer Retention"
author: "Curtis L. Hampton"
output: html_notebook
---

```{r, warning=FALSE}
library(ggplot2)
library(survival)
library(corrplot)
library(car)
```

# Load Data
```{r}
# read data
df <- read.csv("Acquisition and Defection Data.csv")
```

```{r}
# show the structure of the data
str(df)
```

```{r}
# convert variables to factors
cols <- c("Acquisition", "Retention", "Industry")
df[cols] <- lapply(df[cols], factor)
```

```{r}
# create variable names
levels(df$Acquisition) <- make.names(c("No", "Yes"))
levels(df$Retention) <- make.names(c("No", "Yes"))
levels(df$Industry) <- make.names(c("No", "Yes"))
```

```{r}
# delete X column from df
df$X <- NULL
```

```{r}
# show sample of the data
head(df)
```

# ACQUISTION MODELING
Understanding Acquisition: In this first analysis, develop an acquisition model (acquired or not) based upon the following variables: a) acquisition expense, b) Industry, c) Firm revenue and d) Firm employees. Give a basic assessment of the predictive ability of your model and then provide a brief overview of the factors influencing successful acquisition. Since the number of predictor variables is limited, be sure to explore non-linear and/or interaction/moderating effects where possible.

## Logistic Model
```{r}
# fit inital logstic regression model to predict acquisition
acq_glm_fit <- glm(Acquisition ~ Acq_Expense + Industry + Revenue + Employees,
                   data = df, 
                   family = binomial(link = "logit"))
```

```{r}
# fit just the intercept model
int_fit <- glm(Acquisition ~ 1,
               data = df, 
               family = binomial(link = "logit"))
```

```{r}
# show the results of the model
summary(acq_glm_fit)
```

## Check Scatterplot Correlations
```{r}
# subset dataframe to only continuous variables
cont_vars <- df[c("Acq_Expense", "Revenue", "Employees")]
```

```{r}
# create function to provide correlations
panel.cor <- function(x, y, digits = 2, cex.cor, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  # correlation coefficient
  r <- cor(x, y)
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste("r= ", txt, sep = "")
  text(0.5, 0.6, txt)

  # p-value calculation
  p <- cor.test(x, y)$p.value
  txt2 <- format(c(p, 0.123456789), digits = digits)[1]
  txt2 <- paste("p= ", txt2, sep = "")
  if(p<0.01) txt2 <- paste("p= ", "<0.01", sep = "")
  text(0.5, 0.4, txt2)
}
```

```{r}
# create scatterplots and correlations
pairs(cont_vars, upper.panel = panel.cor)
```

## Inspect Model Fit
```{r}
# show VIF for model
vif(acq_glm_fit)
```

```{r}
# check LR test
anova(int_fit,acq_glm_fit,test="LRT")
```

# CHURN MODELING
Modeling Defection: The next step is to model the defection/retention process for those 292 customers which were acquired. In doing so, utilize two different techniques: logistic regression and survival analysis. You may employ any of the variables you feel appropriate in either or both models. Again, alternative expressions of the variables may be useful. Also, can information from the acquisition model be incorporated

## Number of Acquired Customers
```{r}
# number of potential customers vs acquired
table(df$Acquisition)
```

## Customer Lifetime Value vs. Rentention
```{r}
# subset dataset into acquired customers
cust_acq <- df[df$Acquisition == "Yes", ]
```

```{r}
# sort df by Future_CLV
cust_acq <- cust_acq[order(cust_acq$Future_CLV),]
```

```{r}
# plot FutureCLV by Retention
ggplot(data = cust_acq) +
  geom_point(mapping = aes(x = seq(1, length(cust_acq$Future_CLV)), y = Future_CLV, color = Retention)) + 
  ggtitle("Future CLV vs. Rention for Acquired Customers") +
  labs(x = "Rank",y = "Future CLV") +
  theme(plot.title = element_text(lineheight = 0.8, face = "bold", hjust = 0.5)) +
  theme(legend.position = c(0.1, 0.85))
```

## Duration vs. Retention
```{r}
# sort df by Duration
cust_acq <- cust_acq[order(cust_acq$Duration),]
```

```{r}
# plot Duration by Retention
ggplot(data = cust_acq) +
  geom_point(mapping = aes(x = seq(1, length(cust_acq$Duration)), y = Duration, color = Retention)) + 
  ggtitle("Duration vs. Rention for Acquired Customers") +
  labs(x = "Rank",y = "Duration") +
  theme(plot.title = element_text(lineheight = 0.8, face = "bold", hjust = 0.5)) +
  theme(legend.position = c(0.1, 0.85))
```

## Logistic Model
```{r}
# fit inital logstic regression model to predict churn without Future_CLV or Duration
churn_glm_fit <- glm(Retention ~ First_Purchase + Acq_Expense + Ret_Expense + Industry + Revenue + Employees + Breadth + Frequency,
                     data = cust_acq, 
                     family = binomial(link = "logit"))
```

```{r}
# show the results of the model
summary(churn_glm_fit)
```

```{r}
# show VIF for model
vif(churn_glm_fit)
```

## Check Scatterplot Correlations
```{r}
# subset dataframe to only continuous variables
cont_vars <- cust_acq[c("First_Purchase", "Acq_Expense", "Ret_Expense", "Revenue", "Employees", "Breadth", "Frequency")]
```

```{r}
# create scatterplots and correlations
pairs(cont_vars, upper.panel = panel.cor)
```

```{r}
# fit inital logstic regression model to predict churn without Acq_Expense
churn_glm_fit1 <- glm(Retention ~ First_Purchase + Ret_Expense + Industry + Revenue + Employees + Breadth + Frequency,
                      data = cust_acq, 
                      family = binomial(link = "logit"))
```

```{r}
# show the results of the model
summary(churn_glm_fit1)
```

```{r}
# show VIF for model
vif(churn_glm_fit1)
```

### Check Scatterplot Correlations
```{r}
# subset dataframe to only continuous variables
cont_vars <- cust_acq[c("First_Purchase", "Ret_Expense", "Revenue", "Employees", "Breadth", "Frequency")]
```

```{r}
# create scatterplots and correlations
pairs(cont_vars, upper.panel = panel.cor)
```

```{r}
# fit inital logstic regression model to predict churn without First_Purchase
churn_glm_fit2 <- glm(Retention ~ Ret_Expense + Industry + Revenue + Employees + Breadth + Frequency,
                      data = cust_acq, 
                      family = binomial(link = "logit"))
```

```{r}
# show the results of the model
summary(churn_glm_fit2)
```

```{r}
# show VIF for model
vif(churn_glm_fit2)
```

## Check Scatterplot Correlations
```{r}
# subset dataframe to only continuous variables
cont_vars <- cust_acq[c("Ret_Expense", "Revenue", "Employees", "Breadth", "Frequency")]
```

```{r}
# create scatterplots and correlations
pairs(cont_vars, upper.panel = panel.cor)
```

# Survival Analysis
```{r}
# create a "survival object" for each observation, using time and churn data.
cust_acq$survival <- Surv(cust_acq$Duration, cust_acq$Retention == "No")
```

```{r}
# fit a basic survival curve using the data
fit <- survfit(survival ~ 1, data = cust_acq)
```

```{r}
# plot the survival curve and add a title!
plot(fit, lty = 1, mark.time = FALSE, ylim=c(.75,1), xlab = 'Days since Subscribing', ylab = 'Percent Surviving')
```

```{r}
# fit Cox Proportional Hazard Model
churn_coxph <- coxph(survival ~ Ret_Expense + Industry + Revenue + Employees + Breadth + Frequency,
                     data = cust_acq)
```

```{r}
# show the results of the model
summary(churn_coxph)
```